---
title: "Varun_MSiA400_Assignment2"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
redwine <- read.csv("redwine.txt", sep = "\t", header = T)
```


## Problem 1
Mean of RS disregarding missing values
```{r}
mean(redwine$RS, na.rm = T)
```

Similarly, mean of SD disregarding missing values
```{r}
mean(redwine$SD, na.rm = T)
```


## Problem 2

```{r}
SD.obs <- redwine$SD[!is.na(redwine$SD)]
FS.obs <- redwine$FS[!is.na(redwine$SD)]
ABC <- lm(SD.obs ~ FS.obs)
coefficients(ABC)
```
The model is therefore $\hat{SD.obs}$ = 13.185505 + 2.086077 FS.obs

```{r}
summary(ABC)
```
A quick look at the summary of this model shows us that the coefficients are highly significant, indicating a definite linear relationship.

## Problem 3

```{r}
SD.predicted <- predict(ABC, newdata = data.frame(FS.obs = redwine$FS[is.na(redwine$SD)]))
redwine$SD[is.na(redwine$SD)] <- SD.predicted
mean(redwine$SD)
```
The mean of SD after imputation is 46.30182

## Problem 4

```{r}

avg.imp <- function(a,  avg){
missing <- is.na(a)
n.missing <- sum(missing)
a.obs <- a[!missing]
imputed <- a
imputed[missing]<- avg
return(imputed)
}
rsavg <- mean(redwine$RS[!is.na(redwine$RS)])
redwine$RS <- avg.imp(redwine$RS, rsavg)
mean(redwine$RS)
```
Obviously the mean remains unchanged, at 2.53792

## Problem 5

```{r}
fullmodel <- lm(QA~., data = redwine)
coefficients(fullmodel)
```
The coefficients of the full linear model are seen above.

##Problem 6
```{r}
summary(fullmodel)
```

Based on the p-values, the attribute PH is least likely to be related to QA given its large 0.414 p-value.

## Problem 7
```{r}
library(DAAG)
```

```{r}
validation <- CVlm(m = 5, form.lm = formula(QA ~ .), data = redwine)

```

The average error rate for 5-fold validation was seen to be a mean-square error of 63.5.

## Problem 8

```{r}
phmean <- mean(redwine$PH[!is.na(redwine$PH)])
phsd <- sd(redwine$PH[!is.na(redwine$PH)])
ph.ub <- phmean + (3*phsd)
ph.lb <- phmean - (3*phsd)
```

```{r}
redwine2 <- subset(redwine, PH < ph.ub & PH > ph.lb)
nrow(redwine2)
```

redwine2 has 1580 rows, which means 19 observations were removed due to outliers in the attribute PH.


## Problem 9
```{r}
winemodel2 <- lm(QA ~., data = redwine2)
summary(winemodel2)
```
The R squared in the new model has gone up from .358 to .363, suggesting an improvement. Therefore the new model has slightly better performance.

Based on the p-values, the 5 attributes that are most likely to be related to QA now are:
VA, CH, SD, SU and AL, since they have the lowest p-values.
